Full-day, in home validation of infant body position measurements from inertial sensors

### 2023-03-07 manuscript.Rmd
reliability numbers
updated refs

A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video. Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code. Overall agreement averaged XX.X% across video files, ranging from XX-XX. Cohen's *kappa* averaged XX.X% across video files, ranging from XX-XX.

@@

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification [@FranchakScott2021; @NamPark2013, @AiraksinenRasanen2020]. The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right hip X linear acceleration, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for > 75% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package in R [@LiawWiener2002]. Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The *tidymodels* set of packages in R was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns used and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.
<!-- cite tidymodels -->

>>>>

A primary coder completed annotation for the full length of the video, while an independent reliability coder completed annotation for the first thirty minutes of each video. Interrater reliability was based on the proportion of video frames that the two coders chose the same body position code. Overall agreement averaged 90.5% across video files, ranging from 68.4%-100%. Cohen's *kappa* averaged 85.5% across video files, ranging from 31.0%-100%.
<!-- UPDATE THESE BEFORE SUBMITTING -->

@@

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification [@FranchakScott2021; @NamPark2013; @AiraksinenRasanen2020]. The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right hip X linear acceleration, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for > 75% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a "training" set and another, independent portion of the data were defined as a "testing" set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models [@Breiman2001] were trained to predict the human coded body position label for each window from the set of 436 motion features using the *randomForest* package in R [@LiawWiener2002]. Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The *tidymodels* set of packages in R [@tidymodels] was used to tune the "mtry" and "ntrees" hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the *predict* function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

### 2023-03-07 manuscript.tex
updated refs

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification Airaksinen et al. (2020). The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right hip X linear acceleration, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package in R (Liaw et al., 2002). Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The \emph{tidymodels} set of packages in R was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns used and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

>>>>

Classification training and prediction was conducted on a windowed dataset that summarized the raw, 62.5 Hz motion signals within 4-s windows. Overlapping moving windows captured 4-s (250 samples) of data starting each second, which is a common unit of analysis in prior studies of human activity classification (Airaksinen et al., 2020; Franchak et al., 2021; Nam \& Park, 2013). The 250 samples of data in each window were aggregated into a variety of motion features---summary statistics that could be fed into the machine learning model. The minimum, maximum, 25th percentile, 75th percentile, mean, median, skew, kurtosis, standard deviation, and sum were computed for each signal (e.g., right hip X linear acceleration, left ankle pitch angular acceleration). The 10 summary statistics and 24 sensor signals generated 240 columns of motion features that described movement within each window. Furthermore, a series of cross-sensor and cross-orientation summaries (such as the correlation, magnitude, and difference between pairs of sensors) added an additional 196 columns of motion features. The 436 total motion features corresponded to a single body annotation code for each 4-s window. Windows were only used for training/testing if they contained a single body position for \textgreater{} 75\% (3 s) of time within the window to ensure that motion signals could be tied to a clear example of each behavior.

Windowed datasets were used for machine learning classification and validation. For each analysis reported in the results, a subset of data were defined as a ``training'' set and another, independent portion of the data were defined as a ``testing'' set. Different training/testing set combinations were used to address different goals; here we describe the general procedure of how a classifier was derived from a training set. Random forest models (Breiman, 2001) were trained to predict the human coded body position label for each window from the set of 436 motion features using the \emph{randomForest} package in R (Liaw et al., 2002). Random forest models use random subsets of features and random subsets of data across many iterations; this random subsetting creates models that are less likely to be overfit. The \emph{tidymodels} set of packages in R (Kuhn \& Wickham, 2020) was used to tune the ``mtry'' and ``ntrees'' hyperparameters, which control the number of columns sampled and the total number of trees created when modeling. We use an mtry parameter of 44 and and ntree parameter of 550, however, changing hyperparameters had little impact on model performance. Just like a linear regression model, the resulting random forest model could later be applied to a set of testing data with the \emph{predict} function. Validation analyses below will compare the human-coded annotations and random forest model predictions for testing data that are independent from training data, providing a test of whether models can successfully generalize.

@@

\leavevmode\vadjust pre{\hypertarget{ref-tidymodels}{}}%
Kuhn, M., \& Wickham, H. (2020). Tidymodels: A collection of packages for modeling and machine learning using tidyverse principles. \emph{Boston, MA, USA}.
